Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-02-07 Multiresolution_framework/ltx/manuscript.tex
added fig12 in pdf format
loads of minor edits from start to spatial freq section

Typically, the experimentalist only has access the net field generated by the multi-resolution architecture. Follow this, it is natural to ask how do we best describe the multi-resolution structure of the neural system. The solution to this problem was proposed by \citet{Breakspear2005} in using a wavelet decomposition. This paper is directed at developing a framework for patient-specific multi-resolution neural field models, forming a fundamental link between data and theory.

@@

The single layer homogeneous neural field model in the most general form is described by the equation
\begin{equation}
	\label{FullDoubleIntModel} \underbrace{v\left(\mathbf{r},t\right)}_{\text{field}} =
	\int_{-\infty}^t 
	\underbrace{h\left(t - t'\right)}_{\text{synapse}} \int_\Omega
	\parham{\theta}\underbrace{w\left(\mathbf{r},\mathbf{r}'\right)}_{\text{connectivity}} 
	\underbrace{f\left( v\left( \mathbf{r}',t' \right)\right)}_{\text{firing}}
	\, \mathrm{d}\mathbf{r}'\mathrm{d}t',
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ within the spatial domain $\Omega$ is denoted by $v\left(\mathbf r,t\right)$. The function $h(t)$ describes the post-synaptic response kernel, $w(\mathbf{r})$ describes the connectivity kernel \parham{in equation 2 its written $w(r,r')$} and $f(v(\mathbf{r},t))$ describes the relationship between the mean membrane potential and the mean firing rate. The parameter $\theta$ can be thought of the synaptic efficiency or coupling strength, where $\theta>0$ and $\theta<0$ describes excitation and inhibition, respectively. There are various forms of $h(t)$ that have been proposed in the literature, with different levels of complexity. These are shown in Fig.~\ref{fig:Modelcomponents}. Similarly, both the connectivity kernel and the activity function (firing rate) can take various forms. These are also shown in Fig.~\ref{fig:Modelcomponents}. The form of $w(\mathbf{r})$ typically describes an Laplacian (exponential) or Gaussian decay in connectivity strength from any point in the field. By assuming the form of $h(t)$ is the same across layers (scales) the Amari style neural field model is formed, incorporating a multilayer structure into the connectivity kernel. In this way, the connectivity kernel typically takes the form of a Mexican hat (Gaussian shape within layer connectivity) or Wizard hat (Laplacian shape within layer) as depicted in Fig.~\ref{fig:Modelcomponents}.

The stochastic IDE form of the Amari neural field formulation~\cite{Amari1977} is given by (see~\cite{Freestone2011} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, \mathrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where in this first-order (in time) form of the model the membrane dynamics are included through the parameter $\xi=1-T_s/\tau$, $\tau$ is the membrane time constant and $T_s$ is the sampling time step. 
% The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the connectivity kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977,Atay2005}. The components of the model are shown in Fig.~\ref{fig:Modelcomponents}
\begin{figure}[!t]
\centering
\includegraphics[scale=1]{./Graph/fig2.pdf}
\caption{ {\bf The IDE neural field model components}. (\textbf{A}) The Mexican hat (dotted line) and Wizard hat (solid line) connectivity kernels. (\textbf{B}) First order (dotted line) and second order (solid line) synaptic kernels. (\textbf{C}) A representative of the activation function.}
\label{fig:Modelcomponents}
\end{figure}

The term $e_t(\mathbf r)$ is a zero mean Gaussian disturbance, spatially colored but temporally white, with covariance function 
\begin{equation}
cov\left(e_{t}\left(\mathbf{r}\right),e_{t+t'}\left(\mathbf{r'}\right)\right)=\sigma_d\eta(\mathbf{r}-\mathbf{r'})\delta(t-t'),
\label{eq:FieldDisturbance}
\end{equation}
where \parham{$\eta(\mathbf{r}-\mathbf{r'})$ is a spatially homogeneous covariance function, $\sigma_d$ is disturbance variance} and $\delta(\cdot)$ is the Dirac-delta function. %$ = \varsigma v_t(\mathbf{r})$ \cite{VanRotterdam1982,Murphy2009}.

@@

\section{MRA of the IDE in State-Space}
The purpose of forming the multi-resolution approximation (MRA) of the integro-difference equation (IDE) is to form a multi-resolution finite-dimensional state-space model that best captures the spatiotemporal characteristics of the neural field. Given such a model, it is then possible to perform state and parameter estimation in an efficient and robust manner, thus allowing for subject-specific models.

The goal of this section is to demonstrate how to formulate a state-space model in a canonical form such that standard algorithms may be applied to estimate the multi-resolution connectivity structure.   form} a parameterized equation for coefficients of the field basis functions $\mathbf{x}_t$, which becomes the state vector.

The multi-resolution approximation \citep{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(\cdot)$, and the connectivity kernel, $w(\cdot)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering a  one-dimensional field, the static connectivity kernel is decomposed as,
\begin{equation}
 w\left(r\right)=\sum_{l\in \mathbb{Z}}\alpha_{j_0,l} \underbrace{\phi_{j_0,l}\left(r\right)}_{\text{scaling functions}} + \sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l} \underbrace{\psi_{j,l}\left(r\right)}_{\text{wavelets}}, 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with
\begin{align}
\phi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\phi\left(2^j r-l\right) \label{eq:generalscalingfinction}\\	
\psi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\psi\left(2^j r-l\right).  \label{eq:generalwaveletfinction}
\end{align}
The parameters $j$ and $l$ control the scale (dilation) and translation (spatial shift), respectively. \dean{NEED SOMETHING ABOUT HOW j AND l ARE SET HERE - AS IN THAT THEY ARE KNOWN - AND THAT WE ONLY NEED TO ESTIMATE THE COEFFICIENTS.} \parham{The number of spatial shifts, $l$, is dependent on the scale, $j$, and can be chosen such that the basis functions cover the spatial domain of interest. The number of scales depends on the spectral characteristics of the field, which will be described later.} Scaling functions retain the lowest frequency components, due to their low-pass properties, while the wavelet functions extract successively higher frequency components due to their band-pass characteristics.

@@

with covariance (see \citep{Freestone2011} for proof)

@@

\parham{The state-transition equation can be simplified if the chosen family of scaling and wavelet functions are isotropic where}

@@

	\boldsymbol{\lambda}_{i}(r-r') = \boldsymbol{\lambda}(2\mathbf{c}_{i}+r'-r). 

@@

where $c_i$ is the centre of the $i$th basis function of the kernel decomposition. Following this, we can switch the order of operations with the inner product and convolution in Eq.~\eqref{eq:ApproxDiscreteTimeModel4}. With this switch the convolution is now between the scaling functions and wavelets for the field and the scaling functions and wavelets for the connectivity kernel. This convolution can be calculated analytically when using B-spline scaling and wavelet functions and can be predefined once, since these are static functions. This has major computation advantages when solving the estimation problem. \parham{Although this switch holds theoretically it should be noted that it will introduce numerical error, specially close to the edges, due to the finite extent of the field. } The predefined convolution can be written as

@@

 where $\left[\Gamma(r')\right]_{:i}$ denotes the $i$-th column of  $\left[\Gamma(r')\right]$ giving

@@

B-spline wavelet and scaling functions were formulated independently by \citet{Chui1992b,Chui1992,Unser1993}.  The scaling functions are $m$-th order B-splines and the compact support wavelet functions are a linear combination of scaling functions. B-spline wavelet and scaling functions approach optimal time/frequency localization as the order of the spline increases. In fact, for the cubic B-spline scaling and wavelet functions it is already close to the optimal limit for Gaussian functions \citep{Unser1999}. 
%and they have a better approximation rate than other wavelets with the same number of vanishing moments

@@

The $m$-th  order cardinal B-spline function is defined by the recurrence relation \citep{Chui1992} 
\begin{equation}
N_{m}\left(r\right) = \left(N_{m-1}\ast N_{1}\right)\left(r\right) = \int_0^{1} N_{m-1}\left( r-r'\right)\,\mathrm{d}r',
\label{SplineConvolutionIntegral}
\end{equation}
where $m>1$, $\ast$ denotes convolution, and $N_1\left(r\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$, also known as indicator function, i.e.
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
Following this, a B-spline of any order, $N_m(r)$, can be computed using the recursive expression defined by \citep{DeBoor2001}
\begin{equation}\label{eq:MRA-DoBoorFormula}
 N_{m}\left(r\right)=\frac{r}{m-1}N_{m-1}\left(r\right)+\frac{m-r}{m-1}N_{m-1}\left(r-1\right) \quad m>1.
 \end{equation}
The most commonly used form of splines is the cubic spline $\left(m=4\right)$, comprised of third degree polynomials added together at the joining points, with an explicit expression given by
\begin{align}
3!N_{4}\left(r\right)=
\begin{cases}
r^3 & \text{if $ 0\le r\le1$}, \\
4-12r+12r^2-3r^3 & \text{if $1\le r\le2$}, \\
-44+60r-24r^2+3r^3 & \text{if $2\le r\le3$}, \\
64-48r+12r^2-r^3 & \text{if $3\le r\le4$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{align}
One important feature of the B-spline functions is the \emph{partition of unity} property, meaning that the unity can be expressed as a linear sum of B-splines, i.e
\begin{equation}
	\sum_{l}N_m(r+l)\equiv1
	\end{equation}
	\parham{This is important as the underlying field can be reconstructed when its saturated without introducing numerical errors which is not the case for example when Gaussian basis functions are used.}

The convolution and inner product of B-spline functions are required in order to formulate the multi-resolution framework described in the previous section (construction of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Gamma$). To show how the convolution of two B-splines is calculated, Eq.~\eqref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself
\begin{equation}\label{eq:N1convolutions}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution, we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as the support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e. $ N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right)$ (see \ref{ap:InnerProductOfBsplines} for explicit derivation). The values of $N_{m+m'}$ in Eq.~\eqref{eq:BsplineInnerProduct} can be easily determined recursively by evaluating Eq.~\eqref{eq:MRA-DoBoorFormula} at integer points, i.e.
 \begin{equation}\label{eq:MRA-recursiveBsplineatintegerpoints}
 \begin{cases}
 N_2(k)=\delta(k-1)\quad k\in \mathbb{Z}, \\
 N_{m}\left(k\right)=\frac{k}{m-1}N_{m-1}\left(k\right)+\frac{m-k}{m-1}N_{m-1}\left(k-1\right) \quad k=1,2,\dots,m-1
  \end{cases}
 \end{equation}
Note $N_{m}\left(k\right)=0$ for $k\le0$ or $k\ge m$.

The multi-resolution approximation using B-spline functions is completed by defining a two-scale relation pair, which relates the scaling functions and the wavelets at a given scale with the scaling function at the next higher scale. This is important for constructing the wavelets. For cardinal B-spline scaling and wavelet functions of order $m$ the two-scale relation pair takes the form of
\begin{align}
 N_{m}\left(r\right)&=\sum_{n=0}^{m} \left[\mathbf p\right]_n N_{m}\left(2r-n\right) \label{eq:MRA-TwoScalepair1} \\
  \varphi_{m}\left(r\right) &= \sum_{n=0}^{3m-2}  \left[\mathbf q\right]_n N_{m}\left(2r-n\right)\label{eq:MRA-TwoScalepair2},
 \end{align}
where 
 \begin{align}
\left[\mathbf p\right]_n&=2^{-m+1} \binom{m}{n} \quad \text{ $0\le n\le m$} \label{eq:MRA-TwoScalepair1coefs}\\
\left[\mathbf q\right]_n&= \left(-1\right)^n{2^{-m+1}}\sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$}\label{eq:MRA-TwoScalepair2coefs}
 \end{align}
and $\left[\cdot\right]_n$ denotes the $n$-th element of the vector coefficients \citep{Chui1992}. \dean{Need to explicitly say why this is important and what this gives us here.}

Fig.~\ref{fig:MRA-Figure1} shows the cubic B-spline scaling function and its associated wavelet function with dilation and translation parameters set to zero. Note in general the supports of B-spline scaling and wavelet functions are given by 
\begin{align}
	\mathrm{supp}(N_{m;j,l})&=\left[\frac{l}{2^j},\frac{m+l}{2^j}\right] \label{eq:ScalingSupport}\\  
  \mathrm{supp}(\varphi_{m;j,l})&=\left[\frac{l}{2^j},\frac{2m-1+l}{2^j}\right],\label{eq:WaveletSupport}
	\end{align}    
\parham{where subscripts $j$ and $l$ in Eqs.~\eqref{eq:ScalingSupport} and \eqref{eq:WaveletSupport} are introduced to $N_{m}(r)$ and $\varphi_{m}(r)$ to indicate the scale and translation parameters as defined in Eqs~\eqref{eq:generalscalingfinction} and \eqref{eq:generalwaveletfinction}.}
\begin{figure}[!t]
\centering
\includegraphics{./Graph/fig4.pdf}
% \includegraphics{./Graph/Figure1b.pdf}
\caption{{\bf Examples of B-spline scaling and wavelet functions}. (\textbf{A}) The cubic B-spline scaling function. (\textbf{B}) The corresponding wavelet function.}
\label{fig:MRA-Figure1}
\end{figure}

By exploiting Eq.~\eqref{eq:BsplineConvolution} and Eq.~\eqref{eq:BsplineInnerProduct}, the integrals in Eq.~\eqref{eq:Lambdax} and Eq.~\eqref{eq:Gammaij} can be computed analytically. In constructing $\boldsymbol\Lambda_{x}$, it is important to note that B-spline scaling and wavelet functions possess the following orthogonality properties \citep{Unser1993}: 
\begin{equation}
  \left\langle \varphi_{m;j_1,l_1}(r),\varphi_{m;j_2,l_2}(r)\right\rangle =0  \quad \mathrm{for} \quad j_1\neq j_2
 \label{eq:MRA-PsiPsiOrthogonality} 
 \end{equation}
 \begin{equation}
  \left\langle N_{m;j_1,l_1}(r),\varphi_{m;j_2,l_2}(r)\right\rangle =0  \quad \mathrm{for} \quad j_1\leq j_2.
 \label{eq:MRA-PhiPsiOrthogonality}
 \end{equation}
  \parham{Note from Eq.~\eqref{eq:MRA-PsiPsiOrthogonality} B-spline wavelets are semi-orthogonal as they are not orthogonal with respect to translation on a given scale.}

>>>>

Typically, the experimentalist only has access the net field generated by the multi-resolution architecture of the cortex. Follow this, it is natural to ask how do we best describe the multi-resolution structure of the neural system. The solution to this problem was proposed by \citet{Breakspear2005} in using a wavelet decomposition. This paper is directed at developing a framework for patient-specific multi-resolution neural field models, forming a fundamental link between data and theory.

@@

The single layer homogeneous neural field model, in the most general form, is described by the equation
\begin{equation}
	\label{FullDoubleIntModel} \underbrace{v\left(\mathbf{r},t\right)}_{\text{field}} =
	\int_{-\infty}^t 
	\underbrace{h\left(t - t'\right)}_{\text{synapse}} \int_\Omega
	\parham{\theta}\underbrace{w\left(\mathbf{r},\mathbf{r}'\right)}_{\text{connectivity}} 
	\underbrace{f\left( v\left( \mathbf{r}',t' \right)\right)}_{\text{firing}}
	\, \mathrm{d}\mathbf{r}'\mathrm{d}t',
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ within the spatial domain $\Omega$ is denoted by $v\left(\mathbf r,t\right)$. The function $h(t)$ describes the post-synaptic response kernel, $w\left(\mathbf{r},\mathbf{r}'\right)$ describes the connectivity kernel and $f(v(\mathbf{r},t))$ describes the relationship between the mean membrane potential and the mean firing rate. The parameter $\theta$ can be thought of the synaptic efficiency or coupling strength, where $\theta>0$ and $\theta<0$ describes excitation and inhibition, respectively. There are various forms of $h(t)$ that have been proposed in the literature, with different levels of complexity. These are shown in Fig.~\ref{fig:Modelcomponents}. Similarly, both the connectivity kernel and the activity function (firing rate) can take various forms. These are also shown in Fig.~\ref{fig:Modelcomponents}. The form of $w(\mathbf{r},\mathbf{r}')$ typically describes an Laplacian (exponential) or Gaussian decay in connectivity strength from any point in the field. By assuming the form of $h(t)$ is the same across layers (scales) the Amari style neural field model is formed, incorporating a multilayer structure into the connectivity kernel. In this way, the connectivity kernel typically takes the form of a Mexican hat (Gaussian shape within layer connectivity) or Wizard hat (Laplacian shape within layer) as depicted in Fig.~\ref{fig:Modelcomponents}.

The stochastic IDE form of the Amari neural field formulation~\cite{Amari1977} is given by (see~\cite{Freestone2011} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, \mathrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where in this first-order (in time) form of the model the membrane dynamics are included through the parameter $\xi=1-T_s/\tau$, $\tau$ is the membrane time constant and $T_s$ is the sampling time step. 
% The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the connectivity kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977,Atay2005}. The components of the model are shown in Fig.~\ref{fig:Modelcomponents}
\begin{figure}[!t]
\centering
\includegraphics[scale=1]{./Graph/fig2.pdf}
\caption{ {\bf The IDE neural field model components}. (\textbf{A}) The Mexican hat (dotted line) and Wizard hat (solid line) connectivity kernels. (\textbf{B}) First order (dotted line) and second order (solid line) synaptic kernels. (\textbf{C}) A representative of the activation function.}
\label{fig:Modelcomponents}
\end{figure}
The term $e_t(\mathbf r)$ is a zero mean Gaussian disturbance, spatially colored but temporally white, with covariance function 
\begin{equation}
cov\left(e_{t}\left(\mathbf{r}\right),e_{t+t'}\left(\mathbf{r'}\right)\right)=\sigma_d\eta(\mathbf{r}-\mathbf{r'})\delta(t-t'),
\label{eq:FieldDisturbance}
\end{equation}
where \parham{$\eta(\mathbf{r}-\mathbf{r'})$ is a spatially homogeneous covariance function, $\sigma_d$ is disturbance variance} and $\delta(\cdot)$ is the Dirac-delta function. %$ = \varsigma v_t(\mathbf{r})$ \cite{VanRotterdam1982,Murphy2009}.

@@

\section{MRA of the IDE in State-Space}
The purpose of forming the multi-resolution approximation (MRA) of the integro-difference equation (IDE) is to form a multi-resolution finite-dimensional state-space model that best captures the spatiotemporal characteristics of the neural field. Given such a model, it is then possible to perform state and parameter estimation in an efficient and robust manner, thus allowing for subject-specific models. The goal of this section is to demonstrate how to formulate a state-space model in a canonical form such that standard algorithms may be applied to estimate the multi-resolution connectivity structure.   
% form a parameterized equation for coefficients of the field basis functions $\mathbf{x}_t$, which becomes the state vector.

The multi-resolution approximation \citep{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(\cdot)$, and the connectivity kernel, $w(\cdot)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering a  one-dimensional field, the static connectivity kernel is decomposed as,
\begin{equation}
 w\left(r\right)=\sum_{l\in \mathbb{Z}}\alpha_{j_0,l} \underbrace{\phi_{j_0,l}\left(r\right)}_{\text{scaling functions}} + \sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l} \underbrace{\psi_{j,l}\left(r\right)}_{\text{wavelets}}, 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with
\begin{align}
\phi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\phi\left(2^j r-l\right) \label{eq:generalscalingfinction}\\	
\psi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\psi\left(2^j r-l\right).  \label{eq:generalwaveletfinction}
\end{align}
The parameters $j$ and $l$ control the scale (dilation) and translation (spatial shift), respectively. The number of spatial shifts, $l$, is dependent on the scale, $j$, and can be chosen such that the basis functions cover the spatial domain of interest. For example, at scales corresponding to higher spatial frequencies a greater number of wavelets, and thus spatial shifts are required to adequately sample the spatial characteristics of the field. The number of scales depends on the spectral characteristics of the field, where higher spatial frequencies require a more scales. Spatial frequency analysis will be described later in Section~\ref{sec:freq_anal} in greater detail. Scaling functions retain the lowest frequency components, due to their low-pass properties, while the wavelet functions extract successively higher frequency components due to their band-pass characteristics.

@@

with covariance (see \citet{Freestone2011} for proof)

@@

The state-transition equation can be simplified if the chosen family of scaling and wavelet functions are isotropic where

@@

	\boldsymbol{\lambda}_{i}(r-r') = \boldsymbol{\lambda}(2\mathbf{c}_{i}+r'-r), 

@@

where $c_i$ is the centre of the $i^{th}$ basis function of the kernel decomposition. Following this, we can switch the order of operations with the inner product and convolution in Eq.~\eqref{eq:ApproxDiscreteTimeModel4}. With this switch the convolution is now between the scaling functions and wavelets for the field and the scaling functions and wavelets for the connectivity kernel. This convolution can be calculated analytically when using B-spline scaling and wavelet functions and can be predefined, since these are static functions. This has major computation advantages when solving the estimation problem. %\parham{Although this switch holds theoretically it should be noted that it will introduce numerical error, specially close to the edges, due to the finite extent of the field. } 
The predefined convolution can be written as

@@

 where $\left[\Gamma(r')\right]_{:i}$ denotes the $i^{th}$ column of  $\left[\Gamma(r')\right]$ giving

@@

B-spline wavelet and scaling functions were formulated independently by \citet{Chui1992b}, \citet{Chui1992} and \citet{Unser1993}.  The scaling functions are $m$-th order B-splines and the compact support wavelet functions are a linear combination of scaling functions. B-spline wavelet and scaling functions approach optimal time/frequency localization as the order of the spline increases. In fact, for the cubic B-spline scaling and wavelet functions it is already close to the optimal limit for Gaussian functions \citep{Unser1999}. 
%and they have a better approximation rate than other wavelets with the same number of vanishing moments

@@

The $m$-th  order cardinal B-spline function is defined by the recurrence relation \citep{Chui1992} 
\begin{equation}
N_{m}\left(r\right) = \left(N_{m-1}\ast N_{1}\right)\left(r\right) = \int_0^{1} N_{m-1}\left( r-r'\right)\,\mathrm{d}r',
\label{SplineConvolutionIntegral}
\end{equation}
where $m>1$, $\ast$ denotes convolution, and $N_1\left(r\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$, also known as indicator function, i.e.
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
Following this, a B-spline of any order, $N_m(r)$, can be computed using the recursive expression defined by \citet{DeBoor2001}
\begin{equation}\label{eq:MRA-DoBoorFormula}
 N_{m}\left(r\right)=\frac{r}{m-1}N_{m-1}\left(r\right)+\frac{m-r}{m-1}N_{m-1}\left(r-1\right) \quad m>1.
 \end{equation}
The most commonly used form of splines is the cubic spline $\left(m=4\right)$, comprised of third degree polynomials added together at the joining points, with an explicit expression given by
\begin{align}
3!N_{4}\left(r\right)=
\begin{cases}
r^3 & \text{if $ 0\le r\le1$}, \\
4-12r+12r^2-3r^3 & \text{if $1\le r\le2$}, \\
-44+60r-24r^2+3r^3 & \text{if $2\le r\le3$}, \\
64-48r+12r^2-r^3 & \text{if $3\le r\le4$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{align}
One important feature of the B-spline functions is the \emph{partition of unity} property, meaning that the unity can be expressed as a linear sum of B-splines, i.e
\begin{equation}
	\sum_{l}N_m(r+l)\equiv1.
	\end{equation}
This is important as the underlying field can be reconstructed when its saturated without introducing numerical errors which is not the case for example when Gaussian basis functions are used.

The convolution and inner product of B-spline functions are required in order to formulate the multi-resolution framework described in the previous section (construction of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Gamma$). To show how the convolution of two B-splines is calculated, Eq.~\eqref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself
\begin{equation}\label{eq:N1convolutions}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution, we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as the support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e. $ N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right)$ (see \ref{ap:InnerProductOfBsplines} for explicit derivation). The values of $N_{m+m'}$ in Eq.~\eqref{eq:BsplineInnerProduct} can be easily determined recursively by evaluating Eq.~\eqref{eq:MRA-DoBoorFormula} at integer points, i.e.
 \begin{equation}\label{eq:MRA-recursiveBsplineatintegerpoints}
 \begin{cases}
 N_2(k)=\delta(k-1)\quad k\in \mathbb{Z}, \\
 N_{m}\left(k\right)=\frac{k}{m-1}N_{m-1}\left(k\right)+\frac{m-k}{m-1}N_{m-1}\left(k-1\right) \quad k=1,2,\dots,m-1.
  \end{cases}
 \end{equation}
Note $N_{m}\left(k\right)=0$ for $k\le0$ or $k\ge m$.

The multi-resolution approximation using B-spline functions is completed by defining a two-scale relation pair, which relates the scaling functions and the wavelets at a given scale with the scaling function at the next higher scale. This is important for constructing the wavelets. For cardinal B-spline scaling and wavelet functions of order $m$ the two-scale relation pair takes the form of
\begin{align}
 N_{m}\left(r\right)&=\sum_{n=0}^{m} \left[\mathbf p\right]_n N_{m}\left(2r-n\right) \label{eq:MRA-TwoScalepair1} \\
  \varphi_{m}\left(r\right) &= \sum_{n=0}^{3m-2}  \left[\mathbf q\right]_n N_{m}\left(2r-n\right)\label{eq:MRA-TwoScalepair2},
 \end{align}
where 
 \begin{align}
\left[\mathbf p\right]_n&=2^{-m+1} \binom{m}{n} \quad \text{ $0\le n\le m$} \label{eq:MRA-TwoScalepair1coefs}\\
\left[\mathbf q\right]_n&= \left(-1\right)^n{2^{-m+1}}\sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$}\label{eq:MRA-TwoScalepair2coefs}
 \end{align}
and $\left[\cdot\right]_n$ denotes the $n^{th}$ element of the vector coefficients \citep{Chui1992}.

Fig.~\ref{fig:MRA-Figure1} shows the cubic B-spline scaling function and its associated wavelet function with dilation and translation parameters set to zero. Note in general the supports of B-spline scaling and wavelet functions are given by 
\begin{align}
	\mathrm{supp}(N_{m;j,l})&=\left[\frac{l}{2^j},\frac{m+l}{2^j}\right] \label{eq:ScalingSupport}\\  
  \mathrm{supp}(\varphi_{m;j,l})&=\left[\frac{l}{2^j},\frac{2m-1+l}{2^j}\right],\label{eq:WaveletSupport}
	\end{align}    
where subscripts $j$ and $l$ in Eqs.~\eqref{eq:ScalingSupport} and \eqref{eq:WaveletSupport} are introduced to $N_{m}(r)$ and $\varphi_{m}(r)$ to indicate the scale and translation parameters as defined in Eqs~\eqref{eq:generalscalingfinction} and \eqref{eq:generalwaveletfinction}.
\begin{figure}[!t]
\centering
\includegraphics{./Graph/fig4.pdf}
% \includegraphics{./Graph/Figure1b.pdf}
\caption{{\bf Examples of B-spline scaling and wavelet functions}. (\textbf{A}) The cubic B-spline scaling function. (\textbf{B}) The corresponding wavelet function.}
\label{fig:MRA-Figure1}
\end{figure}

By exploiting Eq.~\eqref{eq:BsplineConvolution} and Eq.~\eqref{eq:BsplineInnerProduct}, the integrals in Eq.~\eqref{eq:Lambdax} and Eq.~\eqref{eq:Gammaij} can be computed analytically. In constructing $\boldsymbol\Lambda_{x}$, it is important to note that B-spline scaling and wavelet functions possess the following orthogonality properties \citep{Unser1993}: 
\begin{equation}
  \left\langle \varphi_{m;j_1,l_1}(r),\varphi_{m;j_2,l_2}(r)\right\rangle =0  \quad \mathrm{for} \quad j_1\neq j_2
 \label{eq:MRA-PsiPsiOrthogonality} 
 \end{equation}
 \begin{equation}
  \left\langle N_{m;j_1,l_1}(r),\varphi_{m;j_2,l_2}(r)\right\rangle =0  \quad \mathrm{for} \quad j_1\leq j_2.
 \label{eq:MRA-PhiPsiOrthogonality}
 \end{equation}
Note from Eq.~\eqref{eq:MRA-PsiPsiOrthogonality} B-spline wavelets are semi-orthogonal as they are not orthogonal with respect to translation on a given scale.

In order to analytically calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Gamma$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale using the two-scale relation pair defined in equations (\ref{eq:MRA-TwoScalepair1}-\ref{eq:MRA-TwoScalepair2coefs}). For scaling functions this can be done  by using Eq.~\eqref{eq:MRA-TwoScalepair1} recursively, giving
